---
title: "Create mixture prior"
author: "Yuxin Zou"
date: "12/16/2020"
output: 
  workflowr::wflow_html:
    code_folding: hide
---

# Artificial structure

## 20 traits

Canonical patterns of sharing:
```{r}
R = 20
prior = mmbr:::create_cov_canonical(R)
```

Paired sharing:
```{r}
paired = matrix(0,R,R)
paired[1:2,1:2] = 1
prior[['paired_1']] = paired
paired = matrix(0,R,R)
paired[10:11,10:11] = 1
prior[['paired_2']] = paired
```

Block sharing:
```{r}
block = matrix(0,R,R)
block[1:R/2, 1:R/2] = 1
block[(R/2+1):R, (R/2+1):R] = 1
prior[['blocked_1']] = block
```

We assign weights to priors:

1. singleton total 25%

    * singleton_1 has 15%
    
    * singleton_2 to singleton_11 has 10% (1% each)
    
2. shared total 25%

3. paired 30% (15% each)

4. blocked 20%

```{r}
w = c(0.15, rep(0.01, 10), rep(0, 9), rep(0.05, 5), rep(0.15,2), 0.2)
```

```{r}
prior = prior[which(w>0)]
w = w[which(w>0)]
artificial_mixture_20 = list(U=prior,w=w)
```

## 4 traits

Canonical patterns of sharing:
```{r}
R = 4
prior = mmbr:::create_cov_canonical(R)
```

Paired sharing:
```{r}
paired = matrix(0,R,R)
paired[1:2,1:2] = 1
prior[['paired_1']] = paired
```

Block sharing:
```{r}
block = matrix(0,R,R)
block[1:R/2, 1:R/2] = 1
block[(R/2+1):R, (R/2+1):R] = 1
prior[['blocked_1']] = block
```

We assign weights to priors:

1. singleton total 30%

    * singleton_1, singleton 2 each has 15%

2. shared total 50%

3. paired 10%

4. blocked 10%


```{r}
w = c(0.15, 0.15, 0, 0, rep(0.5/5,5), 0.1, 0.1)
```

```{r}
prior = prior[which(w>0)]
names(prior)
w = w[which(w>0)]
artificial_mixture_4 = list(U=prior,w=w)
```

# Mixture from UKB

The priors are from [this workflow](https://stephenslab.github.io/finemap-uk-biobank/bloodcells_priors.html).

```{r}
prior = readRDS('~/Documents/GitHub/finemap-uk-biobank/output/BloodCells.Ulist.Scor.ed.rds')
names(prior$U)
```
Most weighths are in tFLASH and XX,
```{r}
names(prior$U)[prior$w>0.1]
```
But many other weights are also non-trivial
```{r}
names(prior$U)[prior$w > 0.001]
```
```{r}
tol=1E-12
U = prior$U[which(prior$w>tol)]
w = prior$w[which(prior$w>tol)]
names(U)
bloodcells_mixture = list(U=U,w=w)
```

```{r}
for(i in 1:length(U)){
  bloodcells_mixture$U[[i]] = (bloodcells_mixture$U[[i]] + t(bloodcells_mixture$U[[i]]))/2
  eigenU = eigen(bloodcells_mixture$U[[i]], symmetric = T)
  if(any(eigenU$values<0)){
    eigenU$values[eigenU$values < 0] = 0
    bloodcells_mixture$U[[i]] = eigenU$vectors %*% (t(eigenU$vectors) * eigenU$values)
  }
}
```

```{r}
saveRDS(list(bloodcells_mixture=bloodcells_mixture, artificial_mixture_20=artificial_mixture_20, artificial_mixture_4=artificial_mixture_4), 'output/ukb_prior_simulation.rds')
```
