---
title: "GTEx real data problem continue"
author: "Yuxin Zou"
date: "11/10/2020"
output:
  workflowr::wflow_html:
    code_folding: hide
---

We further investigate the BF for imputed data [here](GTExprofileProblem.html). We fit model with L = 1 and a simple rank 1 prior. We don't estimate prior scalar.

Load data and impute missing data.
```{r}
library(mmbr)
dat = readRDS('data/ENSG00000140265.12.Multi_Tissues.problem.rds')
Ymean = colMeans(dat$Y, na.rm=T)
Y = t(t(dat$Y) - Ymean)
Y[is.na(Y)] = 0
```

Model with dense residual varaince
```{r}
u = rep(1,49)
U = tcrossprod(u)
m = mmbr::msusie(dat$X, Y, prior_variance = U, residual_variance = dat$residual_var, L = 1, compute_objective = T, estimate_prior_variance = FALSE)
susie_plot(m, y='PIP', main=paste0('ELBO=', round(m$elbo[m$niter],2)))
```

Model with diagonal residual variance
```{r}
m_diag = mmbr::msusie(dat$X, Y, prior_variance = U, residual_variance = diag(diag(dat$residual_var)), L = 1, compute_objective = T, estimate_prior_variance = FALSE)
susie_plot(m_diag, y='PIP', main=paste0('ELBO=', round(m_diag$elbo[m_diag$niter],2)))
```

Check CS details:
```{r}
m_diag$sets
```

Suppose there is only one SNP with non-zero effect,
$$
\mathbf{y}_i \sim N_R( x_i \mathbf{b}, V) \quad \mathbf{b} \sim N_R(0, \mathbf{u}\mathbf{u}^\intercal)
$$
Let $z \sim N(0, 1)$, then $\mathbf{b} = z \mathbf{u}$. The model becomes
$$
\mathbf{y}_i \sim N_R( x_i z \mathbf{u}, V) \quad z \sim N(0,1)
$$
The BF for comparing this model with the null model (z = 0) is
$$
BF = (1 + \mathbf{u}^\intercal V^{-1} \mathbf{u} \sum_{i=1}^{N} x_i^2 )^{-1/2} \exp\{\frac{1}{2} \frac{(\mathbf{u}^\intercal V^{-1} \sum_{i=1}^N x_i \mathbf{y}_i)^2}{1 + \mathbf{u}^\intercal V^{-1} \mathbf{u} \sum_{i=1}^{N} x_i^2 } \}
$$

The difference in $V^{-1}$ is
```{r}
summary(as.vector(solve(dat$residual_var) - diag(1/diag(dat$residual_var))))
```

In our example, $u^T V^{-1} u =$ `r round(sum(solve(dat$residual_var)),2)` when V is dense;  $u^T V^{-1} u =$ `r round(sum(1/diag(dat$residual_var)),2)` when V is diagonal. 

```{r fig.width = 30, fig.height = 10}
par(mfrow=c(1,2))
{barplot(colSums(solve(dat$residual_var))/sqrt(sum(colSums(solve(dat$residual_var))^2)), main = 'normalized u^T Vinv, V dense')
barplot((1/diag(dat$residual_var))/sqrt(sum((1/diag(dat$residual_var))^2)), main = 'normalized u^T Vinv, V diagonal')}
```

The difference in results may caused by the difference in the inverse of residual variance matrix.

```{r}
# # BF
# compute_bf = function(X, Y, u, residual_var){
#   J = ncol(X)
#   Y = t(t(Y) - colMeans(Y))
#   X = scale(X)
#   d = colSums(X^2)
#   resid_inv = solve(residual_var)
#   XtY = crossprod(X, Y)
#   lbf = numeric(J)
#   numerator = numeric(J)
#   for(j in 1:J){
#     numerator[j] = (crossprod(u, resid_inv %*% XtY[j,]))^2
#     lbf[j] = - 0.5 * log(1 + crossprod(u, resid_inv %*% u) * d[j]) + 0.5 * ((crossprod(u, resid_inv %*% XtY[j,]))^2) / (1 + crossprod(u, resid_inv %*% u) * d[j] )
#   }
#   return(list(lbf = lbf, numerator = numerator))
# }
# 
# compute_softmax = function(value) {
#     mvalue = max(value)
#     w = exp(value-mvalue)
#     return(as.vector(w / sum(w)))
# }
# 
# lbf_covy = compute_bf(dat$X, Y, u, dat$residual_var)
# lbf_diag = compute_bf(dat$X, Y, u, diag(diag(dat$residual_var)))
# 
# all.equal(compute_softmax(lbf_covy$lbf), m$alpha[1,])
# all.equal(compute_softmax(lbf_diag$lbf), m_diag$alpha[1,])
```

