---
title: "GTEx real data with mmbr commit 72ad394"
author: "Yuxin Zou"
date: "11/5/2020"
output:
  workflowr::wflow_html:
    code_folding: hide
---

In this version of `mmbr`, we simplified some computation, so the speed has improved. We fixed bug in prior scalar estimataion. We implemented ELBO for model with missing data. We computes ELBO for all models and the stopping criteria is based on changes in ELBO. (Without ELBO computation, the stopping criteria is based on changes in pip.)

Here is one gene identified in MASH paper that have different signs for brain vs non brain tissues.

```{r}
# processing code
compute_maf <- function(geno){
  f <- mean(geno,na.rm = TRUE)/2
  return(min(f, 1-f))
}

compute_missing <- function(geno){
  miss <- sum(is.na(geno))/length(geno)
  return(miss)
}

mean_impute <- function(geno){
  f <- apply(geno, 2, function(x) mean(x,na.rm = TRUE))
  for (i in 1:length(f)) geno[,i][which(is.na(geno[,i]))] <- f[i]
  return(geno)
}

is_zero_variance <- function(x) {
  if (length(unique(x))==1) return(T)
  else return(F)
}

filter_X <- function(X, missing_rate_thresh, maf_thresh) {
  rm_col <- which(apply(X, 2, compute_missing) > missing_rate_thresh)
  if (length(rm_col)) X <- X[, -rm_col]
  rm_col <- which(apply(X, 2, compute_maf) < maf_thresh)
  if (length(rm_col)) X <- X[, -rm_col]
  rm_col <- which(apply(X, 2, is_zero_variance))
  if (length(rm_col)) X <- X[, -rm_col]
  return(mean_impute(X))
}

compute_cov_flash <- function(Y, error_cache = NULL){
  covar <- diag(ncol(Y))
  tryCatch({
    fl <- flashier::flash(Y, var.type = 2, prior.family = c(flashier::prior.normal(), flashier::prior.normal.scale.mix()), backfit = TRUE, verbose.lvl=0)
    if(fl$n.factors==0){
      covar <- diag(fl$residuals.sd^2)
    } else {
      fsd <- sapply(fl$fitted.g[[1]], '[[', "sd")
      covar <- diag(fl$residuals.sd^2) + crossprod(t(fl$flash.fit$EF[[2]]) * fsd)
    }
    if (nrow(covar) == 0) {
      covar <- diag(ncol(Y))
      stop("Computed covariance matrix has zero rows")
    }
  }, error = function(e) {
    if (!is.null(error_cache)) {
      saveRDS(list(data=Y, message=warning(e)), error_cache)
      warning("FLASH failed. Using Identity matrix instead.")
      warning(e)
    } else {
      stop(e)
    }
  })
  s <- apply(Y, 2, sd, na.rm=T)
  if (length(s)>1) s = diag(s)
  else s = matrix(s,1,1)
  covar <- s%*%cov2cor(covar)%*%s
  return(covar)
}

get_center <- function(k,n) {
  ## For given number k, get the range k surrounding n/2
  ## but have to make sure it does not go over the bounds
  if (is.null(k)) {
    return(1:n)
  }
  start = floor(n/2 - k/2)
  end = floor(n/2 + k/2)
  if (start<1) start = 1
  if (end>n) end = n
  return(start:end)
}
```

```{r}
dat = readRDS('data/ENSG00000140265.12.Multi_Tissues.rds')
prior = 'data/FastQTLSumStats.mash.FL_PC3.rds'
cis = 500
U = readRDS(prior)$Ulist
weights = rep(1/length(U), length(U))
prior = mmbr::create_mash_prior(mixture_prior=list(weights=weights, matrices=U))
resid_Y = compute_cov_flash(dat$y_res)
X = filter_X(dat$X, 0.1, 0.05)
X = X[,get_center(cis, ncol(X))]
Y = dat$y_res
```

The covariance/correlation matrix of Y using pairwise complete observations:
```{r, fig.width = 30, fig.height = 15}
library(corrplot)
par(mfrow=c(1,2))
corrplot(cov(Y, use='pairwise.complete.obs'), method='color', type='upper', tl.col="black", tl.srt=45, is.corr = FALSE)
corrplot(cor(Y, use='pairwise.complete.obs'), method='color', type='upper', tl.col="black", tl.srt=45, is.corr = TRUE)
```

The covarince/correlation matrix of Y using FLASH:
```{r, fig.width = 30, fig.height = 15}
colnames(resid_Y) = rownames(resid_Y) = colnames(Y)
par(mfrow=c(1,2))
corrplot(resid_Y, method='color', type='upper', tl.col="black", tl.srt=45, is.corr = FALSE)
corrplot(cov2cor(resid_Y), method='color', type='upper', tl.col="black", tl.srt=45, is.corr = TRUE)
```

## Models with L = 10

We fit 3 models with L = 10:

1. model with exact computation;

2. model with approximate computation;

3. model with approximate computation using diagonal residual variance, which is equivalent to exact computation with diagonal residual variance.

```{r, eval=FALSE}
stime <- proc.time()
res <- mmbr::msusie(X, Y, prior_variance=prior, residual_variance=resid_Y, approximate=FALSE, compute_objective = T)
etime <- proc.time()
time_res <- etime - stime
saveRDS(list(result = res, result_time = time_res), 'output/GTExprofile_res_elbo.rds')
rm(res)

stime <- proc.time()
res_approx <- mmbr::msusie(X, Y, prior_variance=prior, residual_variance=resid_Y, approximate=TRUE, compute_objective = T)
etime <- proc.time()
time_res_approx <- etime - stime
saveRDS(list(result = res_approx, result_time = time_res_approx), 'output/GTExprofile_resapprox_elbo.rds')
rm(res_approx)

stime <- proc.time()
res_approx_diag <- mmbr::msusie(X, Y, prior_variance=prior, residual_variance=diag(diag(resid_Y)), approximate=TRUE, compute_objective = T)
etime <- proc.time()
time_res_approx_diag <- etime - stime
saveRDS(list(result = res_approx_diag, result_time = time_res_approx_diag), 'output/GTExprofile_resapproxdiag_elbo.rds')
rm(res_approx_diag)
```

Load models:
```{r}
library(mmbr)
res1 = readRDS('output/GTExprofile_res_elbo.rds')
res2 = readRDS('output/GTExprofile_resapprox_elbo.rds')
res3 = readRDS('output/GTExprofile_resapproxdiag_elbo.rds')
```

Model 1 credible sets:
```{r}
susie_plot(res1$result, y='PIP', main=paste0('dense V, exact, ELBO=',round(res1$result$elbo[res1$result$niter], 2)))
```
Model 2 credible sets:
```{r}
susie_plot(res2$result, y='PIP', main=paste0('dense V, approx, ELBO=', round(res2$result$elbo[res2$result$niter], 2)))
```
Model 3 credible sets:
```{r}
susie_plot(res3$result, y='PIP', main=paste0('diagonal V, ELBO=', round(res3$result$elbo[res3$result$niter], 2)))
```

There is no overlapping between CSs. The blue CS in model 3 does not included in CSs from model 1 and 2. 


|         | Total Time                      |  Algorithm Time                     | # iterations          
|---------|---------------------------------|-------------------------------------|-----------------------
| model 1 | `r res1$result_time['elapsed']` | `r res1$result$walltime['elapsed']` | `r res1$result$niter` 
| model 2 | `r res2$result_time['elapsed']` | `r res2$result$walltime['elapsed']` | `r res2$result$niter`
| model 3 | `r res3$result_time['elapsed']` | `r res3$result$walltime['elapsed']` | `r res3$result$niter`

[Univariate Effects](assets/GRExProfile/GTExprofile_univ.pdf)

```{r, eval=FALSE}
p = mmbr::mmbr_plot(res1$result)
pdf('docs/assets/GRExProfile/GTExprofile_res_elbo.pdf', width = 60, height = 15)
print(p$plot)
dev.off()
```
[Model 1 Effects](assets/GRExProfile/GTExprofile_res_elbo.pdf)

```{r, eval=FALSE}
p = mmbr::mmbr_plot(res2$result)
pdf('docs/assets/GRExProfile/GTExprofile_resapprox_elbo.pdf', width = 60, height = 15)
print(p$plot)
dev.off()
```
[Model 3 Effects](assets/GRExProfile/GTExprofile_resapprox_elbo.pdf)

```{r, eval=FALSE}
p = mmbr::mmbr_plot(res3$result)
pdf('docs/assets/GRExProfile/GTExprofile_resapproxdiag_elbo.pdf', width = 5, height = 15)
print(p$plot)
dev.off()
```
[Model 4 result](assets/GRExProfile/GTExprofile_resapproxdiag_elbo.pdf)

## Models with L = 3

```{r, eval=FALSE}
stime <- proc.time()
res <- mmbr::msusie(X, Y, prior_variance=prior, residual_variance=resid_Y, approximate=FALSE, L=3, compute_objective = T)
etime <- proc.time()
time_res <- etime - stime
saveRDS(list(result = res, result_time = time_res), 'output/GTExprofile_resL3_elbo.rds')

stime <- proc.time()
res_approx <- mmbr::msusie(X, Y, prior_variance=prior, residual_variance=resid_Y, approximate=TRUE, L=3, compute_objective = T)
etime <- proc.time()
time_res_approx <- etime - stime
saveRDS(list(result = res_approx, result_time = time_res_approx), 'output/GTExprofile_resapproxL3_elbo.rds')

stime <- proc.time()
res_approx_diag <- mmbr::msusie(X, Y, prior_variance=prior, residual_variance=diag(diag(resid_Y)), approximate=TRUE, L=3, compute_objective = T)
etime <- proc.time()
time_res_approx_diag <- etime - stime
saveRDS(list(result = res_approx_diag, result_time = time_res_approx_diag), 'output/GTExprofile_resapproxdiagL3_elbo.rds')
```

```{r}
res1_L3 = readRDS('output/GTExprofile_resL3_elbo.rds')
res2_L3 = readRDS('output/GTExprofile_resapproxL3_elbo.rds')
res3_L3 = readRDS('output/GTExprofile_resapproxdiagL3_elbo.rds')
```

Model 1 credible sets:
```{r}
susie_plot(res1_L3$result, y='PIP', main=paste0('dense V, exact, ELBO=', round(res1_L3$result$elbo[res1_L3$result$niter], 2)))
```
Model 2 credible sets:
```{r}
susie_plot(res2_L3$result, y='PIP', main=paste0('dense V, approx, ELBO=', round(res2_L3$result$elbo[res2_L3$result$niter], 2)))
```
Model 3 credible sets:
```{r}
susie_plot(res3_L3$result, y='PIP', main=paste0('diagonal V, ELBO=', round(res3_L3$result$elbo[res3_L3$result$niter], 2)))
```

|         | Total Time                      |  Algorithm Time                     | # iterations          
|---------|---------------------------------|-------------------------------------|-----------------------
| model 1 | `r res1_L3$result_time['elapsed']` | `r res1_L3$result$walltime['elapsed']` | `r res1_L3$result$niter`
| model 2 | `r res2_L3$result_time['elapsed']` | `r res2_L3$result$walltime['elapsed']` | `r res2_L3$result$niter`
| model 3 | `r res3_L3$result_time['elapsed']` | `r res3_L3$result$walltime['elapsed']` | `r res3_L3$result$niter`

```{r, eval=FALSE}
p = mmbr::mmbr_plot(res1_L3$result)
pdf('docs/assets/GRExProfile/GTExprofile_resL3_elbo.pdf', width = 17, height = 15)
print(p$plot)
dev.off()
```
[Model 1 L3 Effects](assets/GRExProfile/GTExprofile_resL3_elbo.pdf)

## Models with L = 1

```{r, eval=FALSE}
stime <- proc.time()
res <- mmbr::msusie(X, Y, prior_variance=prior, residual_variance=resid_Y, approximate=FALSE, L=1, compute_objective = T)
etime <- proc.time()
time_res <- etime - stime
saveRDS(list(result = res, result_time = time_res), 'output/GTExprofile_resL1_elbo.rds')

stime <- proc.time()
res_approx <- mmbr::msusie(X, Y, prior_variance=prior, residual_variance=resid_Y, approximate=TRUE, L=1, compute_objective = T)
etime <- proc.time()
time_res_approx <- etime - stime
saveRDS(list(result = res_approx, result_time = time_res_approx), 'output/GTExprofile_resapproxL1_elbo.rds')

stime <- proc.time()
res_approx_diag <- mmbr::msusie(X, Y, prior_variance=prior, residual_variance=diag(diag(resid_Y)), approximate=TRUE, L=1, compute_objective = T)
etime <- proc.time()
time_res_approx_diag <- etime - stime
saveRDS(list(result = res_approx_diag, result_time = time_res_approx_diag), 'output/GTExprofile_resapproxdiagL1_elbo.rds')
```

```{r}
res1_L1 = readRDS('output/GTExprofile_resL1_elbo.rds')
res2_L1 = readRDS('output/GTExprofile_resapproxL1_elbo.rds')
res3_L1 = readRDS('output/GTExprofile_resapproxdiagL1_elbo.rds')
```

Model 1 credible sets:
```{r}
susie_plot(res1_L1$result, y='PIP', main=paste0('dense V, exact, ELBO=', round(res1_L1$result$elbo[res1_L1$result$niter], 2)))
```
Model 2 credible sets:
```{r}
susie_plot(res2_L1$result, y='PIP', main=paste0('dense V, approx, ELBO=', round(res2_L1$result$elbo[res2_L1$result$niter], 2)))
```
Model 3 credible sets:
```{r}
susie_plot(res3_L1$result, y='PIP', main=paste0('diagonal V, ELBO=', round(res3_L1$result$elbo[res3_L1$result$niter], 2)))
```

The CS from model 3 does not overlap any CSs from model 1.
