---
title: "GTEx real data"
author: "Yuxin Zou"
date: "10/19/2020"
output:
  workflowr::wflow_html:
    code_folding: hide
---

Here is one gene that identified in MASH paper that seems to have different direction of effects in brain vs non-brain tissues.

```{r}
# processing code
compute_maf <- function(geno){
  f <- mean(geno,na.rm = TRUE)/2
  return(min(f, 1-f))
}

compute_missing <- function(geno){
  miss <- sum(is.na(geno))/length(geno)
  return(miss)
}

mean_impute <- function(geno){
  f <- apply(geno, 2, function(x) mean(x,na.rm = TRUE))
  for (i in 1:length(f)) geno[,i][which(is.na(geno[,i]))] <- f[i]
  return(geno)
}

is_zero_variance <- function(x) {
  if (length(unique(x))==1) return(T)
  else return(F)
}

filter_X <- function(X, missing_rate_thresh, maf_thresh) {
  rm_col <- which(apply(X, 2, compute_missing) > missing_rate_thresh)
  if (length(rm_col)) X <- X[, -rm_col]
  rm_col <- which(apply(X, 2, compute_maf) < maf_thresh)
  if (length(rm_col)) X <- X[, -rm_col]
  rm_col <- which(apply(X, 2, is_zero_variance))
  if (length(rm_col)) X <- X[, -rm_col]
  return(mean_impute(X))
}

compute_cov_flash <- function(Y, error_cache = NULL){
  covar <- diag(ncol(Y))
  tryCatch({
    fl <- flashier::flash(Y, var.type = 2, prior.family = c(flashier::prior.normal(), flashier::prior.normal.scale.mix()), backfit = TRUE, verbose.lvl=0)
    if(fl$n.factors==0){
      covar <- diag(fl$residuals.sd^2)
    } else {
      fsd <- sapply(fl$fitted.g[[1]], '[[', "sd")
      covar <- diag(fl$residuals.sd^2) + crossprod(t(fl$flash.fit$EF[[2]]) * fsd)
    }
    if (nrow(covar) == 0) {
      covar <- diag(ncol(Y))
      stop("Computed covariance matrix has zero rows")
    }
  }, error = function(e) {
    if (!is.null(error_cache)) {
      saveRDS(list(data=Y, message=warning(e)), error_cache)
      warning("FLASH failed. Using Identity matrix instead.")
      warning(e)
    } else {
      stop(e)
    }
  })
  s <- apply(Y, 2, sd, na.rm=T)
  if (length(s)>1) s = diag(s)
  else s = matrix(s,1,1)
  covar <- s%*%cov2cor(covar)%*%s
  return(covar)
}

get_center <- function(k,n) {
  ## For given number k, get the range k surrounding n/2
  ## but have to make sure it does not go over the bounds
  if (is.null(k)) {
    return(1:n)
  }
  start = floor(n/2 - k/2)
  end = floor(n/2 + k/2)
  if (start<1) start = 1
  if (end>n) end = n
  return(start:end)
}
```

```{r}
dat = readRDS('data/ENSG00000140265.12.Multi_Tissues.rds')
prior = 'data/FastQTLSumStats.mash.FL_PC3.rds'
cis = 500
U = readRDS(prior)$Ulist
weights = rep(1/length(U), length(U))
prior = mmbr::create_mash_prior(mixture_prior=list(weights=weights, matrices=U))
resid_Y = compute_cov_flash(dat$y_res)
X = filter_X(dat$X, 0.1, 0.05)
X = X[,get_center(cis, ncol(X))]
Y = dat$y_res
```

We fit 4 models:

1. model with exact computation

2. model with exact computation using diagonal residual variance

3. model with approximate computation

4. model with approximate computation using diagonal residual variance

We expect model 2 and 4 have same results.

```{r, eval=FALSE}
library(profvis)
prof1 = profvis({
  stime <- proc.time()
  res <- mmbr::msusie(X, Y, prior_variance=prior, residual_variance=resid_Y, approximate=FALSE)
  etime <- proc.time()
  time_res <- etime - stime
  },prof_output='output/mmbr_gtex_res.Rprof')
saveRDS(list(result = res, result_time = time_res, result_profile = prof1), 'output/GTExprofile_res.rds')
rm(res)
rm(prof1)

prof2 = profvis({
  stime <- proc.time()
  res_diag <- mmbr::msusie(X, Y, prior_variance=prior, residual_variance=diag(diag(resid_Y)), approximate=FALSE)
  etime <- proc.time()
  time_res_diag <- etime - stime
  },prof_output='output/mmbr_gtex_res_diag.Rprof')
saveRDS(list(result = res_diag, result_time = time_res_diag, result_profile = prof2), 'output/GTExprofile_resdiag.rds')
rm(res_diag)
rm(prof2)

prof3 = profvis({
  stime <- proc.time()
  res_approx <- mmbr::msusie(X, Y, prior_variance=prior, residual_variance=resid_Y, approximate=TRUE)
  etime <- proc.time()
  time_res_approx <- etime - stime
  },prof_output='output/mmbr_gtex_res_approx.Rprof')
saveRDS(list(result = res_approx, result_time = time_res_approx, result_profile = prof3), 'output/GTExprofile_resapprox.rds')
rm(res_approx)
rm(prof3)

prof4 = profvis({
  stime <- proc.time()
  res_approx_diag <- mmbr::msusie(X, Y, prior_variance=prior, residual_variance=diag(diag(resid_Y)), approximate=TRUE)
  etime <- proc.time()
  time_res_approx_diag <- etime - stime
  },prof_output='output/mmbr_gtex_res_approx_diag.Rprof')
saveRDS(list(result = res_approx_diag, result_time = time_res_approx_diag, result_profile = prof4), 'output/GTExprofile_resapproxdiag.rds')
rm(res_approx_diag)
rm(prof4)
```

Load models:
```{r}
library(mmbr)
res1 = readRDS('output/GTExprofile_res.rds')
res2 = readRDS('output/GTExprofile_resdiag.rds')
res3 = readRDS('output/GTExprofile_resapprox.rds')
res4 = readRDS('output/GTExprofile_resapproxdiag.rds')
```

As we expected, model 2 and 4 are same.
```{r}
all.equal(res2$result$pip, res4$result$pip)
```

We check results from model 1, 3 and 4.

Model 1 credible sets:
```{r}
susie_plot(res1$result, y='PIP')
```
Model 3 credible sets:
```{r}
susie_plot(res1$result, y='PIP')
```
Model 4 credible sets:
```{r}
susie_plot(res4$result, y='PIP')
```

Model 1 and 3 gives same Credible Sets. 

|         | Total Time                      |  Algorithm Time                     | # iterations          |
|---------|---------------------------------|-------------------------------------|-----------------------|
| model 1 | `r res1$result_time['elapsed']` | `r res1$result$walltime['elapsed']` | `r res1$result$niter` |
| model 2 | `r res2$result_time['elapsed']` | `r res2$result$walltime['elapsed']` | `r res1$result$niter` |
| model 3 | `r res3$result_time['elapsed']` | `r res3$result$walltime['elapsed']` | `r res1$result$niter` |
| model 4 | `r res4$result_time['elapsed']` | `r res4$result$walltime['elapsed']` | `r res1$result$niter` |


```{r, eval=FALSE}
univariate_res = lapply(1:ncol(Y), function(i) susieR:::univariate_regression(X,Y[,i]))
res1$result$bhat = do.call(cbind, lapply(1:ncol(Y), function(i) univariate_res[[i]]$betahat))
res1$result$shat = do.call(cbind, lapply(1:ncol(Y), function(i) univariate_res[[i]]$sebetahat))
rownames(res1$result$bhat) = 1:501
p = mmbr::mmbr_plot(res1$result, original_sumstat = TRUE, cs_only = FALSE)
pdf('docs/assets/GRExProfile/GTExprofile_univ.pdf', width = 100, height = 15)
print(p$plot)
dev.off()
```
[Univariate Effects](assets/GRExProfile/GTExprofile_univ.pdf)

```{r, eval=FALSE}
p = mmbr::mmbr_plot(res1$result)
pdf('docs/assets/GRExProfile/GTExprofile_res.pdf', width = 60, height = 15)
print(p$plot)
dev.off()
```
[Model 1 Effects](assets/GRExProfile/GTExprofile_res.pdf)

```{r, eval=FALSE}
p = mmbr::mmbr_plot(res3$result)
pdf('docs/assets/GRExProfile/GTExprofile_resapprox.pdf', width = 60, height = 15)
print(p$plot)
dev.off()
```
[Model 3 Effects](assets/GRExProfile/GTExprofile_resapprox.pdf)

```{r, eval=FALSE}
p = mmbr::mmbr_plot(res4$result)
pdf('docs/assets/GRExProfile/GTExprofile_resapproxdiag.pdf', width = 5, height = 15)
print(p$plot)
dev.off()
```
[Model 4 result](assets/GRExProfile/GTExprofile_resapproxdiag.pdf)
