---
title: "M&M RSS V Problem"
author: "Yuxin Zou"
date: "02/28/2020"
output:
  workflowr::wflow_html:
    code_folding: hide
---

```{r}
library(mmbr)
DSC = dscrutils:::load_inputs(c('data/lite_data_6.rds','data/lite_data_6_shared_1.pkl', 'data/lite_data_6_shared_1_get_sumstats_1.rds', 'data/lite_data_6_shared_1_get_sumstats_1_oracle_generator_1_mnm_rss_shared_16.rds'), dscrutils:::read_dsc)
X = DSC$X
meta = DSC$meta
ss = DSC$sumstats
Z = as.matrix(ss$bhat/ss$shat)
r = cor(X)
prior = DSC$prior
Y = DSC$Y_sim
true_pos = as.integer(apply(meta$true_coef, 1, sum) != 0)
```

The simulated data has one causal SNP shared across all 5 conditions.
\begin{align}
Y &= XB + E, \\
B &= \gamma b^\intercal, \\
b &\sim N_5(0, 0.3^2 \left(\begin{matrix}
1 & 0.999 & 0.999 & 0.999 & 0.999 \\
0.999 & 1 & 0.999 & 0.999 & 0.999 \\
0.999 & 0.999 & 1 & 0.999 & 0.999 \\
0.999 & 0.999 & 0.999 & 1 & 0.999 \\
0.999 & 0.999 & 0.999 & 0.999 & 1
\end{matrix}\right)), \\
E &\sim MN(0, I, \left(\begin{matrix} 
0.0366 & 0 & 0 & 0 & 0 \\
0 & 0.0337 & 0 & 0 & 0 \\
0 & 0 & 0.0357 & 0 & 0 \\
0 & 0 & 0 & 0.0351 & 0 \\
0 & 0 & 0 & 0 & 0.036
\end{matrix}\right)),
\end{align}
where $X$ is column standardized.

We compute z scores from simple linear regression, which has following distribution
\begin{align}
\hat{Z} &\sim MN(RZ, R, V), \\
V &= diag(sd(Y)) \left(\begin{matrix} 
0.0366 & 0 & 0 & 0 & 0 \\
0 & 0.0337 & 0 & 0 & 0 \\
0 & 0 & 0.0357 & 0 & 0 \\
0 & 0 & 0 & 0.0351 & 0 \\
0 & 0 & 0 & 0 & 0.036
\end{matrix}\right) diag(sd(Y)) \\
&= \left( \begin{matrix}
0.8977 & 0 & 0 & 0 & 0 \\
0 & 0.896 & 0 & 0 & 0 \\
0 & 0 & 0.9325 & 0 & 0 \\
0 & 0 & 0 & 0.867 & 0 \\
0 & 0 & 0 & 0 & 0.8983
\end{matrix} \right)
\end{align}

## Oracle residual variance

```{r}
resid_Z_oracle <- meta$residual_variance/apply(Y, 2, var)
xUlist = lapply(prior$xUlist, function(U) t(t(U*nrow(Y)/apply(Y, 2, sd))/apply(Y, 2, sd)))
# xUlist = lapply(prior$xUlist, function(U) U*nrow(Y))
m_init = mmbr::create_mash_prior(mixture_prior = list(matrices=xUlist, weights=prior$pi), null_weight=prior$null_weight, max_mixture_len=-1)
fit1 = mmbr::msusie_rss(Z, r, L=1, prior_variance=m_init, residual_variance=resid_Z_oracle, compute_objective=TRUE, estimate_residual_variance=F, estimate_prior_variance=T, estimate_prior_method='simple')
susieR::susie_plot(fit1,y='PIP', xlab = 'SNP positions', b=true_pos)
```

The elbo is
```{r}
fit1$elbo
```

## Estimate V from Z

```{r}
resid_Z_all = cor(Z)
round(resid_Z_all, 4)
```

```{r}
m_init = mmbr::create_mash_prior(mixture_prior = list(matrices=xUlist, weights=prior$pi), null_weight=prior$null_weight, max_mixture_len=-1)
fit2 = mmbr::msusie_rss(Z, r, L=1, prior_variance=m_init, residual_variance=resid_Z_all, compute_objective=TRUE, estimate_residual_variance=F, estimate_prior_variance=T, estimate_prior_method='simple')
susieR::susie_plot(fit2,y='PIP', xlab = 'SNP positions', b=true_pos)
```

The elbo is
```{r}
fit2$elbo
```

## Estimate V from null Z

```{r}
max_absz = apply(abs(Z),1, max)
nullish = which(max_absz < 2) # length 17
nullish_z = Z[nullish,]
resid_Z_null <- cor(nullish_z)
resid_Z_null
```

```{r}
m_init = mmbr::create_mash_prior(mixture_prior = list(matrices=xUlist, weights=prior$pi), null_weight=prior$null_weight, max_mixture_len=-1)
fit3 = mmbr::msusie_rss(Z, r, L=1, prior_variance=m_init, residual_variance=resid_Z_null, compute_objective=TRUE, estimate_residual_variance=F, estimate_prior_variance=T, estimate_prior_method='simple')
susieR::susie_plot(fit3,y='PIP', xlab = 'SNP positions', b=true_pos)
```

The elbo is
```{r}
fit3$elbo
```

## Estimate V as identity

```{r}
resid_Z_identity = diag(ncol(Z))
```

```{r}
m_init = mmbr::create_mash_prior(mixture_prior = list(matrices=xUlist, weights=prior$pi), null_weight=prior$null_weight, max_mixture_len=-1)
fit4 = mmbr::msusie_rss(Z, r, L=1, prior_variance=m_init, residual_variance=resid_Z_identity, compute_objective=TRUE, estimate_residual_variance=F, estimate_prior_variance=T, estimate_prior_method='simple')
susieR::susie_plot(fit4,y='PIP', xlab = 'SNP positions', b=true_pos)
```

The elbo is
```{r}
fit4$elbo
```
