---
title: "RSS problem data 1"
author: "Yuxin Zou"
date: "7/11/2020"
output:
  workflowr::wflow_html:
    code_folding: hide
---

There are 3 causals in this dataset.

```{r}
library(mmbr)
X = readRDS('data/tiny_data_211.rds')
simu = readRDS('data/tiny_data_211_artificial_mixture_small_missing_2.rds')
b = simu$meta$true_coef
Z = simu$sumstats$bhat/simu$sumstats$sbhat
r = X$ld
prior = simu$meta$prior[["oracle"]]
resid_Z <- simu$meta$residual_variance
xUlist = prior$xUlist
m_init = mmbr::create_mash_prior(mixture_prior = list(matrices=xUlist, weights=prior$pi), null_weight=prior$null_weight, max_mixture_len=-1)
```

The total PVE for each condition is
```{r}
0.15 * apply(b, 2, function(x) length(which(x!=0)))
```
Per SNP PVE is 0.15, wihch is a little high, this may violate the RSS model assumption.

The true effects are at `r which(rowSums(b!=0) !=0)`.
```{r}
b[which(rowSums(b!=0) !=0),]
```

The Z scores at causal are
```{r}
Z[which(rowSums(b!=0) !=0),]
```

The model with individual level is
```{r}
result0 = mmbr::msusie(X$X, simu$Y, L = 10, 
                       prior_variance=m_init, residual_variance=resid_Z, 
                       compute_objective=T, estimate_residual_variance=F,
                       estimate_prior_variance=T, estimate_prior_method='EM', 
                       precompute_covariances=T, n_thread=1, max_iter=1000)
susie_plot(result0,y='PIP',b=b)
```

The model using summary data is
```{r}
result = mmbr::msusie_rss(Z, r, L=10, 
                          prior_variance=m_init, residual_variance=resid_Z, 
                          compute_objective=TRUE, estimate_residual_variance=F, 
                          estimate_prior_variance=T, estimate_prior_method='EM', 
                          precompute_covariances=T, n_thread=1, max_iter=1000, track_fit = T)
susie_plot(result,y='PIP',b=b)
```
There is only one CS containing causal, the rest 8 CSs are all false discoveries.

Change residual variance to residual correlation matrix (encourage conservative),
```{r}
result.cor = mmbr::msusie_rss(Z, r, L=10, 
                          prior_variance=m_init, residual_variance=cov2cor(resid_Z), 
                          compute_objective=TRUE, estimate_residual_variance=F, 
                          estimate_prior_variance=T, estimate_prior_method='EM', 
                          precompute_covariances=T, n_thread=1, max_iter=1000, track_fit = T)
susie_plot(result.cor,y='PIP',b=b)
```
None if CS contain signal, but the elbo is higher (`r susie_get_objective(result.cor)` vs `r susie_get_objective(result)`).

Try initialize at truth:
```{r}
init_true = list()
init_true$b1 = array(0, dim = c(3,301,6))
init_true$b1[1,66,] = (b[66,]/sqrt(diag(resid_Z))) * sqrt(837)
init_true$b1[2,118,] = (b[118,]/sqrt(diag(resid_Z))) * sqrt(837)
init_true$b1[3,236,] = (b[236,]/sqrt(diag(resid_Z))) * sqrt(837)
init_true$alpha = matrix(0, 3, 301)
init_true$alpha[1,66] = 1
init_true$alpha[2,118] = 1
init_true$alpha[3,236] = 1
result.init.true = msusie_rss(Z, r, L=3, s_init = init_true,
                              prior_variance=m_init, residual_variance=resid_Z, 
                              compute_objective=TRUE, estimate_residual_variance=F, 
                              estimate_prior_variance=T, estimate_prior_method='EM', 
                              precompute_covariances=T, n_thread=1, max_iter=1000,track_fit=T)
susie_plot(result.init.true, y='PIP',b=b)
```

L = 1:
```{r}
result.1 = mmbr::msusie_rss(Z, r, L=1, 
                            prior_variance=m_init, residual_variance=resid_Z, 
                            compute_objective=TRUE, estimate_residual_variance=F, 
                            estimate_prior_variance=T, estimate_prior_method='EM', 
                            precompute_covariances=T, n_thread=1, max_iter=1000, track_fit = T)
susie_plot(result.1,y='PIP',b=b)
```
The correlation between SNP 30 and causal 236 is `r r[30,236]`.

L = 2:
```{r}
result.2 = mmbr::msusie_rss(Z, r, L=2, 
                            prior_variance=m_init, residual_variance=resid_Z, 
                            compute_objective=TRUE, estimate_residual_variance=F, 
                            estimate_prior_variance=T, estimate_prior_method='EM', 
                            precompute_covariances=T, n_thread=1, max_iter=1000, track_fit = T)
susie_plot(result.2,y='PIP',b=b)
```

L = 3:
```{r}
result.3 = mmbr::msusie_rss(Z, r, L=3, 
                            prior_variance=m_init, residual_variance=resid_Z, 
                            compute_objective=TRUE, estimate_residual_variance=F, 
                            estimate_prior_variance=T, estimate_prior_method='EM', 
                            precompute_covariances=T, n_thread=1, max_iter=1000, track_fit = T)
susie_plot(result.3,y='PIP',b=b)
```

```{r}
result.3$sets
```

Correlation between CS1 and CS3 is
```{r}
r[30, c(29,236,260)]
```

L = 4:
```{r}
result.4 = mmbr::msusie_rss(Z, r, L=4, 
                            prior_variance=m_init, residual_variance=resid_Z, 
                            compute_objective=TRUE, estimate_residual_variance=F, 
                            estimate_prior_variance=T, estimate_prior_method='EM', 
                            precompute_covariances=T, n_thread=1, max_iter=1000, track_fit = T)
susie_plot(result.4,y='PIP',b=b)
```

