---
title: "mvSuSiE naive prior vs random prior"
author: "Yuxin Zou"
date: "06/19/2022"
output:
  workflowr::wflow_html:
    code_folding: hide
---

```{r}
library(mvsusieR)
library(reshape2)
library(ggplot2)
plot_sharing = function(X, col = 'black', to_cor=FALSE, title="", remove_names=F) {
  clrs <- colorRampPalette(rev(c("#D73027","#FC8D59","#FEE090","#FFFFBF",
                                 "#E0F3F8","#91BFDB","#4575B4")))(128)
  if (to_cor) lat <- cov2cor(X)
  else lat = X/max(diag(X))
  lat[lower.tri(lat)] <- NA
  n <- nrow(lat)
  if (remove_names) {
    colnames(lat) = paste('t',1:n, sep = '')
    rownames(lat) = paste('t',1:n, sep = '')
  }
  melted_cormat <- melt(lat[n:1,], na.rm = TRUE)
  melted_cormat$Var2 = as.factor(melted_cormat$Var2)
  melted_cormat$Var1 = as.factor(melted_cormat$Var1)
  p = ggplot(data = melted_cormat, aes(Var2, Var1, fill = value))+
    geom_tile(color = "white")+ggtitle(title) + 
    scale_fill_gradientn(colors = clrs, limit = c(-1,1), space = "Lab") +
    theme_minimal()+ 
    coord_fixed() +
    theme(
      axis.title.x = element_blank(),
      axis.title.y = element_blank(),
      axis.text.x = element_text(color=col, size=8,angle=45,hjust=1),
      axis.text.y = element_text(color=rev(col), size=8),
      title =element_text(size=10),
      # panel.grid.major = element_blank(),
      panel.border = element_blank(),
      panel.background = element_blank(),
      axis.ticks = element_blank(),
      legend.justification = c(1, 0),
      legend.position = c(0.6, 0),
      legend.direction = "horizontal")+
    guides(fill = guide_colorbar(title="", barwidth = 7, barheight = 1,
                                 title.position = "top", title.hjust = 0.5))
  if(remove_names){
    p = p + scale_x_discrete(labels= 1:n) + scale_y_discrete(labels= n:1)
  }
  
  return(p)
}
```

In simulation with UKB blood cell traits priors, the random effect prior performs slightly better than canonical prior.

Load simulated data
```{r}
dat = readRDS('data/analysis_20220619/data_ukb_609_ukb_bloodcells_mixture_2.rds')
meta = dat$meta
```

There are 3 causal variants.

Load the model using canonical prior
```{r}
m_can = readRDS('data/analysis_20220619/data_ukb_9_ukb_bloodcells_mixture_1_mnm_rss_naive_corZ_1.rds')
susie_plot(m_can$result, y='PIP', b = meta$true_coef)
```

Load the model using random effects prior
```{r}
m_ran = readRDS('data/analysis_20220619/data_ukb_9_ukb_bloodcells_mixture_1_mnm_rss_identity_corZ_1.rds')
susie_plot(m_ran$result, y='PIP', b = meta$true_coef)
```

The variant 1375 has PIP 0.78 using canonical prior, PIP 0.99 using random effects prior. The true effect of variant 1375 is
```{r}
rename = list('WBC_count' = 'WBC#',
              'RBC_count' = 'RBC#',
              'Haemoglobin' = 'HGB',
              'MCV' = 'MCV',
              'RDW' = 'RDW',
              "Platelet_count" = 'PLT#',
              "Plateletcrit" = 'PCT',
              "PDW" = 'PDW',
              "Lymphocyte_perc" = 'LYMPH%',
              "Monocyte_perc" = 'MONO%',
              "Neutrophill_perc" = 'NEUT%',
              "Eosinophill_perc" = 'EO%',
              "Basophill_perc" = 'BASO%',
              "Reticulocyte_perc" = 'RET%',
              "MSCV" = 'MSCV', 
              "HLR_perc" = 'HLR%')
trait_names = sapply(colnames(meta$residual_variance), function(x) rename[[x]])
bloodcells_col = cbind(trait_names, 
                       c('Compound white cell', 'Mature red cell', 'Mature red cell', 
                         'Mature red cell', 'Mature red cell', 'Platelet', 'Platelet',
                         'Platelet', 'Compound white cell', 'Compound white cell', 
                         'Compound white cell', 'Compound white cell', 'Compound white cell',
                         'Immature red cell', 'Mature red cell','Immature red cell'),
                       c('#33cccc', 'red', 'red', 'red', 'red',
                         '#cc66ff', '#cc66ff', '#cc66ff',
                         '#33cccc', '#33cccc', '#33cccc', '#33cccc', '#33cccc',
                         'pink', 'red', 'pink'))
trait_new_order = c("RBC#", "HGB", "MCV", "RDW", "MSCV", "RET%", "HLR%", "PLT#", "PCT", "PDW", 
                    "WBC#", "LYMPH%", "MONO%", "NEUT%", "EO%", "BASO%")
traits_index = match(trait_new_order, trait_names)

colnames(meta$true_coef) = colnames(meta$residual_variance)
barplot(meta$true_coef[1375,traits_index], las = 2, cex.names=0.6)
```
It is simulated from this covaraince structure
```{r}
plot_sharing(meta$trueU$`1375`[traits_index, traits_index], col=bloodcells_col[traits_index,3]) 
```

The posterior weights for each component in canonical prior is
```{r}
w = m_can$result$mixture_weights[2, 1375,]
names(w) = c('null', names(meta$prior$naive$xUlist))
round(w, 4)
```
